\center \fontsize{18}{20} \selectfont Revealing the Landscape of Globally Color-Dual Multi-loop Integrands

### 2023-09-27 notes/CSTheory.tex
CS theory edits

\documentclass[11pt,letter]{article}
\pdfoutput=1
\usepackage{jheppub}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{slashed}
\usepackage{anyfontsize}
\usepackage{tocloft}
\renewcommand{\cftdot}{.}
\renewcommand{\cftsubsecleader}{\cftdotfill{\cftdotsep}}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{bbold}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{orcidlink}
\usepackage[normalem]{ulem}

@@

\author{\large Alex Edison \orcidlink{0000-0002-5430-9500},}
\author{\large James Mangan  \orcidlink{0000-0002-9713-7446},}
\author{\large and Nicolas H. Pavao \orcidlink{0000-0002-9817-8266}}

@@

\section{Introduction}\label{intro}
Since the turn of the century, our understanding of the $S$-matrix and
its concealed structures has expanded dramatically. At the heart of this
progress is the mantra that physical observables are simpler when
studied on-shell
\cite{Parke:1986gb,TasiLance,Cheung:2017pzi}. However, while inserting on-shell states certainly offers a path towards taming the
factorial growth of Feynman diagrams, it also obscures the off-shell
simplicity at the heart of many quantum field theories. A shining
example of this hidden structure is the duality between color and
kinematics \cite{BCJ,Bern:2010ue,BCJreview}, which states that the
kinematic numerators of gauge theory amplitudes can be rearranged to
obey the same algebraic identities as the constituent color
factors. When this duality is realized globally off-shell, it
dramatically decreases the combinatorial complexity posed by integrand
construction, and offers a path towards efficient assembly of quantum
gravity integrands directly from simpler gauge theory building blocks
via the double copy \cite{BCJ,Bern:2010ue}.

Despite the tremendous success in leveraging this duality to
compute gauge and gravity observables to high orders in perturbation
theory \cite{FiveLoopN4, GeneralizedDoubleCopyFiveLoops, Bern:2018jmv,
  Carrasco:2021otn}, color-kinematics remains a conjecture at loop
level. Indeed, there are many examples in the literature where
identifying color-dual representations beyond one-loop has posed a
formidable challenge
\cite{Mogull:2015adi, Johansson:2017bfl, KalinN2TwoLoop, Bern:2015ooa}. In this work, we use
the nonlinear sigma model (NLSM) and Yang-Mills, two theories proven
to permit color-dual representations at tree-level
\cite{Feng:2010my,Cachazo:2014xea}, as case studies for advancing our
understanding of the kinematic algebra toward the multi-loop level.

We begin with an overview of color-kinematics duality in
\cref{background}, and define the notion of ``globally'' color-dual integrands in \cref{sec:bootstrap}. 
We then construct a manifestly color-dual theory in
\cref{sec:cubic}, which generates $D$-dimensional color-dual $n$-point
numerators for both NLSM and the self-dual sector of Yang-Mills
through one-loop. The construction of this theory, which we dub
\textit{semi-abelian Yang-Mills}, relies on isolating the cubic sector
of pure Yang-Mills amplitudes from the four-point contacts that are
needed to fully realize non-abelian gauge symmetry. Despite the
potency of this theory at one-loop, the construction runs into an
obstruction at two-loop, which we describe in \cref{2LoopObstruction}.

Faced with this obstruction, we study the multi-loop sector of both
NLSM and Yang-Mills in \cref{2loopBoot} using an ansatz based
color-dual bootstrap. In contrast to much of the available literature
on color-dual integrand construction at two-loop
\cite{Bern:2013yya,Mogull:2015adi,Johansson:2017bfl}, our results are
completely agnostic to the spacetime dimension; all the polarizations
and momenta appearing in our construction will remain formally
$D$-dimensional. This makes our methods and results particularly well
suited for algorithmically extracting rational terms from the
$D$-dependence of dimensionally regulated loop momenta.
The ansatz approach to constructing numerators generally results in an explosion of terms, but in the case of scalar theories there is an added difficulty because the linear equations become dense.
We were able overcome this barrier by employing a custom
solver, \texttt{FiniteFieldSolve}, which will soon be made
public.

With this solver, we are able to compute a two-loop integrand for NLSM
that globally manifests the duality off-shell. This represents a new
benchmark for color-dual representations in non-supersymmetric gauge
theories.  However in sharp contrast, we find that Yang-Mills does
\emph{not} permit a globally color-dual representation, even when
considering the most general polynomial ansatz of Lorentz covariant
kinematics.\footnote{Globally color-dual local numerators for two-loop
  pure YM have been studied before but only in the context of a
  limited ansatz subject to certain loop power counting constraints
  \cite{Bern:2015ooa}.}  Concretely, the failure point can be
pinpointed to a conflict between between the ``bowtie'' cut and the
following Jacobi triple:
\begin{equation}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
We comment on the implications of this finding in \cref{conclusions},
and discuss alternative approaches where one must relax field
theoretic locality in order to accommodate the construction of
globally color-dual integrands at generic loop order.

\section{Background}\label{background}
To set the stage, in this section we provide an overview of
color-kinematics duality -- both the on-shell and off-shell
constructions. We then detail a color-dual bootstrap approach for
constructing integrands in theories for which the fully off-shell kinematic
algebra has not yet been identified, like Yang-Mills and NLSM.
\subsection{Color-dual Amplitudes}\label{onShellCK}

Color-kinematics duality is a statement about the off-shell graphical
simplicity that manifests redundancy in the on-shell observables of
certain field theories. Ignoring factors of $i$ and the coupling
constant, the starting point to realize the duality is to express generic $n$-point $L$-loop amplitudes in a cubic graph representation:
\begin{equation}
  \mathcal{A}_n^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{C_\Gamma N_\Gamma}{d_\Gamma} .
  \label{eq:gen-amp}
\end{equation}
Here the spacetime dimension is $D$.  Even though the underlying
theory may have quartic (or higher multiplicity) vertices, like YM and NLSM, it is
possible to multiply and divide by propagators so that the sum runs
over the set of purely cubic graphs, $\Gamma$.  Overcounting is removed by
the internal symmetry factor $S_\Gamma$, which tracks the number of
graph symmetries with external legs held fixed.

The color factor
$C_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu, Carrasco:2020ywq,Carrasco:2022jxn,
    Carrasco:2023vjg}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $N_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors. In this 
form the color factors $C_\Gamma$ are not independent and are related to
one another via a set of Jacobi identities that look schematically
like
\begin{equation}
C_i+C_j+C_k=0,
\end{equation}
where $i$, $j$, and $k$ are three graphs.  Example Jacobi relations
are presented in detail in \cref{sec:bootstrap}. The duality between color and kinematics is the statement that that there exists a way to rearrange factors between
the various kinematic numerators $N_\Gamma$ such that the numerators
obey the same Jacobi identities as the color factors
\begin{equation}
N_i +N_j +N_k=0.
\end{equation}
If color-kinematics duality can be achieved at the off-shell level,
then the kinematic numerators should satisfy an algebra just like the
color factors.  In theories where the kinematic algebra is known, it
is often the algebra of volume preserving diffeomorphisms
\cite{Monteiro2011pc, Ben-Shahar:2021zww, Cheung:2020djz,
  Cheung:2021zvb, Cheung:2022mix}.
When the kinematic algebra is not known explicitly, the kinematic numerators are typically constructed from an ansatz.
As an example, the result of such a calculation for four-point NLSM scattering at tree level produces the numerators
\begin{align}\label{pionNums}
&N^{\text{NLSM}}_s = s(t-u)\\
&N^{\text{NLSM}}_t = t(u-s)\\
&N^{\text{NLSM}}_u = u(s-t)
\end{align}
where the coupling has been normalized away.  These numerators sum to
zero by explicit calculation.  Note that the $s$-channel numerator
contains an explicit factor of $s$ that cancels with the propagator
and the same goes for the $t$- and $u$-channels.
The net result is that the amplitude is a local function, just as one would expect for the pion tree
amplitude.

Once color-dual numerators $N_\Gamma$ have been found, they can be
used to replace the color factors ($C_\Gamma \to N_\Gamma$) in a
\emph{different} amplitude to produce the corresponding amplitude in
a new theory,
\be\label{DCconstruction}
\tilde{\mathcal{A}}_n^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{C_\Gamma
  \tilde{N}_\Gamma}{d_\Gamma} ~~ \xrightarrow{(C_\Gamma \to N_\Gamma)}
~~ \mathcal{M}_n^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{N_\Gamma
  \tilde{N}_\Gamma}{d_\Gamma}.
\end{equation}
This is known as the double-copy construction \cite{BCJ,Bern:2010ue,BCJreview}.
Importantly, the $\tilde{N}_\Gamma$ do not need to respect
color-kinematics duality.  The prototypical example of the double-copy
is that YM double copied with YM results in gravity where the gauge
invariance of each separate gluon produces the diffeomorphism
invariance of the graviton.  The double copy can be proven at tree
level using the Kawai-Lewellen-Tye (KLT) relations but there is significant evidence that
the double copy persists to all loop orders \cite{FiveLoopN4,
  GeneralizedDoubleCopyFiveLoops, Bern:2018jmv, Carrasco:2021otn, KLT,
  KiermaierTalk}.  In addition to pointing at undiscovered
structure hidden in the Lagrangian for gravity, the double copy is
immensely useful at a utilitarian level since it is much more
efficient at producing $\mathcal{M}_n^L$ than traditional methods like
Feynman rules.

Many of the one-loop results presented in \cref{sec:cubic} are informed
by the tree-level double-copy so we will review the tree case here.
The color structure of tree-level amplitudes is particularly simple.
A fully color dressed tree amplitude $\mathcal{A}_n$ can be decomposed
in terms of the trace basis
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-1}} \text{Tr}(T^{a_{\sigma_1}} T^{a_{\sigma_2}}...T^{a_{\sigma_{n-1}}} T^{a_n}) A_n[\sigma_1, \sigma_2,...\sigma_{n-1}, n],
  \label{eq:trace-basis}
\end{equation}
where $T^a$ are generators of the gauge group and, because of the
cyclicity of the trace, one of the legs is held fixed so that the sum
runs over the $(n-1)!$ permutations of the remaining external legs.
The coefficients of the color factors are the color-ordered partial
amplitudes $A_n[...]$.  Any tree-level color factor can be converted
into a linear combination of Del Duca-Dixon-Maltoni (DDM) half-ladder
color factors
\begin{equation}
\text{DDM}[a_1, a_2,... a_n] \equiv f^{a_1 a_2 b_1} f^{b_1 a_3 b_2} f^{b_2 a_4 b_3}...f^{b_{n-3} a_{n-1} a_n}
\end{equation}
associated with the graph
\begin{equation}
\halfLadder
\end{equation}
by repeated application of the Jacobi identity \cite{DixonMaltoni}.
The fully color-dressed tree amplitude can then be re-expressed as
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-2}} \text{DDM}[a_1, a_{\sigma_2}, a_{\sigma_3}, a_{\sigma_{n-1}}, a_n]  A_n[1, \sigma_2,...\sigma_{n-1}, n]
  \label{eq:ddm-basis}
\end{equation}
with legs $1$ and $n$ fixed.  Since the the sum only contains $(n-2)!$
terms, there must be additional relations -- the
Kleiss-Kuijf (KK) relations \cite{Kleiss:1988ne} -- amongst the
partial amplitudes appearing in the right-hand side of
\cref{eq:trace-basis}.  The KK relations are generic to any theory
with purely adjoint particles.  Color-kinematics duality further
implies that the KK basis is overcomplete and can be reduced to a
basis of $(n-3)!$ amplitudes via the fundamental Bern-Carrasco-Johansson (BCJ) identities
\cite{BCJ,Feng:2010my}
\begin{equation}
\sum \limits_{i=2}^{n-1} k_1 \cdot (k_2+...+k_i) A_n[2,...,i,1,i+1,...,n] =0.
\end{equation}
In terms of constructing explicit color-dual solutions at tree level,
it is enough to specify the numerator of the half-ladder because any
diagram can be reduced to a half-ladder through successive
applications of the Jacobi identity.  At one-loop, every diagram can be
related to one topology as well, in this case the $n$-gon, which can be understood as a forward limit of the half-ladder.  Two-loop
calculations play an important role in our understanding of
color-kinematics duality because this is the first order where
multiple basis graphs are required, at least for a generic theory. We refer the reader to \cite{BCJreview} for more background on the topic.

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to trivialize
the duality by encoding it in an off-shell Lagrangian
of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form
\begin{equation}\label{cubicCKLag}
\mathcal{L} = P_V^{ab}P_W^{\mu\nu}\mathcal{O}^a_\mu \Box \mathcal{O}^b_\nu +  V^{abc} W^{\mu \nu \rho}\mathcal{O}^a_\mu \mathcal{O}^b_\nu \mathcal{O}^c_\rho
\end{equation}
where $V^{abc}$ and $W^{\mu \nu \rho}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_\mu$ indexed by quantum numbers $a$ and
$\mu$. The Feynman rules for this theory are simply:
\be\label{CKFrules}
\begin{split}
\propFrule{$a,\mu$}{$b,\nu$}{$k \rightarrow $}{} &= \frac{i}{k^2} (P_V^{-1})^{ab}(P_W^{-1})^{\mu\nu}
\\
\cubicFrule{$a,\mu$}{$b,\nu$}{$c,\rho$}{}{}{} &= i V^{abc}W^{\mu \nu \rho}
\end{split}
\ee
We have introduced $P_V^{ab}$ and $P_W^{\mu\nu}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction, as is the case for $DF^2$ theory \cite{Johansson:2017srf}. Generally, these quadratic dressings are simply delta functions, $P_V^{ab} = \delta^{ab}$, or flat space metrics, $P_W^{\mu\nu}=\eta^{\mu\nu}$. As written, the theory will be color dual if the off-shell cubic vertex is antisymmetric and satisfies the Jacobi identity. Specifically,
\begin{align}
\text{anti-symmetry}:& \quad{}_{\mu_1}\langle W^{\mu_2} \rangle_{\mu_3}+\text{cyc}(23)=0
\\
\text{Jacobi identity}:& \quad {}_{\mu_1}\langle W^{\mu_2} W^{\mu_3}\rangle_{\mu_4}+\text{cyc}(234)=0\label{jacID}
\end{align}
where, using the Feynman rules of \cref{CKFrules}, the bracketed expression is the half-ladder numerator of the following diagram,
\be
{}_{\mu_1}\langle W^{\mu_2}W^{\mu_3}\cdots W^{\mu_{n-1}} \rangle_{\mu_n} = N\left[\halfLadderKinetic\right]\,,
\ee
and likewise for the $V^{abc}$ dressing.
Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}.
Typically when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional operators \cite{Carrasco:2022lbm,Carrasco:2022sck} or propagating fields \cite{Ben-Shahar:2022ixa}.
Indeed, as we will show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic two-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.

\subsection{Color-dual Bootstrap}
\label{sec:bootstrap}

When an off-shell color-dual Lagrangian is not known, the loop-level
amplitude integrand must be constructed from an ansatz.  The four
conditions we impose on an integrand ansatz are summarized here and
elaborated on below.\footnote{The four ansatz conditions are similar
  to those of Ref.~\cite{Bern:2015ooa} except that there the authors
  demand that the numerators have the same loop power counting as
  Feynman gauge Feynman rules.  This is an additional assumption that
  is most emphatically not made here.}
After imposing the four constraints below, any remaining coefficients in the ansatz encode ``generalized gauge freedom'', meaning that the
choice of coefficients does not affect the physical integrand or its double copy.
\begin{enumerate}
\item \textbf{Off-shell Locality}: Diagram numerators are polynomials
  in momenta and polarizations, e.g., the diagrams only have poles given
  by the propagators of the diagrams.
\item \textbf{Color-kinematics duality}: The numerators obey the same
  Jacobi identities as the color factors.  This implicitly requires
  that the integrand is expressed in terms of purely cubic diagrams.
\item \textbf{Graph symmetries}: A diagram's numerator is only a
  function of the diagram's topology and labeling.  Furthermore, the
  diagram numerators are invariant under the automorphisms of the
  diagrams including signs that compensate for color-factor sign
    changes.
\item \textbf{On-shell Unitarity}: The cuts of the ansatz must
  reproduce the physical unitarity cuts of the theory when internal
  momenta are taken on-shell.
\end{enumerate}

In order to clarify general statements, the following subsections
include examples from a four-point two-loop integrand for some generic
color-dual theory like YM.  This is the simplest case that
demonstrates the full machinery of the color-dual bootstrap.  All tree
and one-loop processes are generated by single basis diagrams (the half
ladder and $n$-gon respectively) so they lack examples of three-term
``boundary'' Jacobi relations to be described shortly.
Choosing the four-point two-loop integrand also has the advantage that this
process appears prominently in this paper.

\paragraph{Off-shell Locality}

The kinematic numerators must be $D$-dimensional, Lorentz invariant,
local (polynomial) functions of the graph kinematics with the correct
power counting and external states. Locality is physically motivated since Feynman
rules produce polynomial numerators. Indeed, the most natural way to guarantee locality and avoid spurious poles is to require the off-shell numerators to be polynomial functions of
kinematics. However, insisting on a local numerator is
partly a matter of convenience as the ansatz for a rational function
would quickly grow out of control without a guiding principle for what
terms to include.  While these
assumptions are well motivated from a physical and complexity
standpoint, the literature has featured numerators that violate
locality, modify naive power counting through the inclusion of higher
spin modes, and break manifest Lorentz invariance \cite{Square,
  WeinzierlBCJLagrangian, Mogull:2015adi, FivePointN4BCJ,
  Johansson:2017bfl, Ben-Shahar:2022ixa, Cheung:2016prv,
  Cheung:2021zvb}.  These developments suggest that relaxing locality
is a viable alternative when the condition proves too stringent for an
ansatz.  However, simply requiring on-shell locality may suffice.
This alternative will be discussed in \cref{nonLocalScattering}.

\paragraph{Color-kinematics duality}

Color-kinematics duality is a statement about color factors, kinematic
numerators, and the accompanying Jacobi relations.  Adjoint-type color
factors are associated with cubic graphs, so satisfying
color-kinematics duality implicitly requires expressing the integrand
in terms of such graphs. As mentioned earlier, this is possible to do
even for a theory with quartic (or higher multiplicity) vertices by
multiplying and dividing by propagators.  In the case of the four-point
two-loop example promised above, there are 14 cubic graph topologies
(shown in \cref{fig:MaxCuts}) ignoring tadpoles and bubble on external
leg (BEL) graphs.

\begin{figure}
\centering
{\MCutC, \MCutD, \sumoCut, \MCutJ,  
 \\
  \doubleBoxCut,  \crossBoxCut,\pentaTriangleCut,  \MCutF, 
\\
\MCutI,  \MCutE, \MCutK, 
\\
 \MCutL, \MCutG,  \MCutH
}
\caption{The 14 non-pathological four-point two-loop cubic diagrams.} \label{fig:MaxCuts}
  \end{figure}

With the integrand expressed in terms of cubic graphs, it is possible
to discuss the main ingredient in color-kinematics duality -- the
Jacobi relations.  The color factors in an integrand obey a set of
Jacobi identities.  Color-kinematics duality states that there
\emph{exists} a way of writing the kinematic numerators of the
integrand so that they obey the same set of Jacobi identities.  The
Jacobi identities can be divided into two categories: \emph{defining}
relations and \emph{boundary} relations.  Defining Jacobi relations
can be used to express any graph in terms of a basis.  For example,
the double box, crossed box, and penta-triangle are related to each
other as follows
\begin{equation}
\label{eq:TwoLoopJacobiOne}
\JacobiOneDoubleBox + \JacobiOneCrossedBox + \JacobiOnePentaTriangle
=0\,,
\end{equation}
where the shared propagator is shown in red.  Using this relation the
crossed box can be written (or \emph{defined}) in terms of the other
two.  In this way, all of the 14 cubic four-point two-loop topologies can be
written in terms of a basis of the double box and penta-triangle.
(Any two of the double box, crossed box, and penta-triangle would
suffice for a basis but the planar basis is chosen for simplicity in
this work.)  The numerator of each of the two basis graphs receives
its own ansatz and the numerator of any other graph can be expressed
in terms of these two.

After using the defining Jacobi relations to relate every graph to the
basis, any remaining Jacobi relations will be referred to as boundary
relations.  In other words, boundary relations (indirectly) relate
basis elements to themselves.  For example, the crossed box numerator
is related to itself via the following Jacobi relation
\begin{equation}
\label{eq:TwoLoopJacobiTwo}
\JacobiTwoCrossedBoxOne + \JacobiTwoCrossedBoxTwo +  \JacobiTwoCrossedBoxThree =0,
\end{equation}
where the shared propagator is shown in red as before.  Each of the
crossed box numerators can be written in terms of the double box and
penta-triangle so this Jacobi relation places constraints on the basis
numerator \ansatze{}.  While defining Jacobi relations simplify how
many numerators require \ansatze{}, boundary relations place
restrictions on those \ansatze{}.  In principle there could be Jacobi
relations involving tadpoles and BEL graphs, which would be subtle to
handle.  One option is to force the numerators of these graphs to
zero.  However, since these graphs will not contribute to physical
unitarity cuts, we take a different approach.  These graphs are
allowed to have non-zero numerators but any Jacobi relations involving
them are simply not imposed or solved.  See Refs.~\cite{Bern:2012uf,
  Edison:2022jln} and \cref{BELreg} for discussions of BELs in the context of
color-kinematics duality.

What we have just described is
sometimes referred to as ``global'' color-kinematics duality because
the Jacobi identities are solved by a single (global) integrand with
off-shell internal lines.  In order to perform the double copy it is
enough to simply satisfy the Jacobi identities on the generalized
unitarity cuts, meaning that each graph can be assigned its own ansatz
\cite{Bern:2015ooa}.  If a global solution cannot be found, this is
one way to enlarge the ansatz, where another option is to incorporate non-local terms in the ``numerators''.  A globally
color-dual integrand is desirable because of its simplicity and
because a color-dual Lagrangian would naively produce precisely this
object.

\paragraph{Graph symmetries}

Every numerator is required to respect all of the symmetries (or
automorphisms) of its associated graph.  Graph symmetries are an
avatar of Bose symmetry.  For a pure ansatz construction it is very
reasonable to impose graph symmetries, but color-kinematics duality at
tree-level is often achieved at the expense of manifest Bose symmetry
\cite{Cheung:2016prv, Cheung:2021zvb, Brandhuber:2021bsf,
  Edison:2020ehu}.  In an ansatz construction graph symmetries are
imposed for \emph{every} graph, not just the basis graphs.
For example, the symmetries of the crossed box are spanned by,
\begin{equation}
\scalebox{0.8}{\CrossedBoxGraphSym{1}{2}{3}{4}}\!\! \sim\!\! \scalebox{0.8}{\CrossedBoxGraphSym{2}{1}{3}{4}}
\text{ and }
\scalebox{0.8}{\CrossedBoxGraphSym{1}{2}{3}{4}} \!\!\sim\!\! \scalebox{0.8}{\CrossedBoxGraphSym{2}{1}{4}{3}}
\end{equation}
which are just functional relabelings of external legs,
$\{1234\} \to \{2134\}$ and $\{1234\} \to \{1243\}$.  In general,
internal edge labels must be tracked as well.  Furthermore, care must
be taken with the signs in a symmetry relation.  To compensate for the
antisymmetric color factors, $f^{abc}$, graph vertices are all totally
antisymmetric and hence have signs built into their orientations.
When used in conjuction with the Jacobi relations, \eg{}
\cref{eq:TwoLoopJacobiOne}, the symmetry constraints of non-basis
graphs still impose conditions on the basis ansatz.  In fact, since
the symmetry properites of a graph are directly linked with
transformations of its color dressing, the resulting constraints can
be thought of as a subclass of the \emph{boundary} Jacobi relations.

\paragraph{On-shell Unitarity}

The final constraints come from ensuring that the ansatz correctly
reproduces the generalized unitarity cuts of the desired theory, thus
guaranteeing that the numerators encode the correct physical,
theory-dependent information.  Generalized unitarity has been
reviewed in many places
\cite{BCFUnitarity, Bern:2015ooa, Bourjaily:2017wjl,
  JJHenrikReview, FivePointN4BCJ, Bern:2012uf, BCJreview,
  Edison:2022jln, Edison:2022smn} and was recently optimized for effective field theories like NLSM \cite{Carrasco:2023qgz}. Generalized unitarity equates products of on-shell tree amplitudes,
encoded via a graph $\gamma$, to sums over compatible\footnote{A
  diagram $g$ is compatible with the cut diagram $\gamma$ if $g$ is
  isomorphic to $\gamma$ or to any of the additional factorization
  channels of $\gamma$. } diagram numerators evaluated on the support
of the on-shell conditions and weighted by their uncut propagators
\begin{equation}
\label{eq:SchematicCutEquation}
\text{Cut}[\gamma]= \sum_{\substack{\text{states}\\\text{crossing } E(\gamma)}} \prod_{v \in V(\gamma)} \atree(v)
= \sum_{g \text{ compat } \gamma}
\frac{N[g] \vert_\text{cut}}{\prod \limits_{\text{uncut}} \ell^2}
 \, ,
\end{equation}
where $V(\gamma)$ are the vertices of $\gamma$ and $E(\gamma)$ its
edges.  The sum over states hides much of the complexity of cut
construction: for scalars the process is trivial while spinning states
greatly complicate matters (see
Refs.~\cite{Kosmopoulos:2020pcd,Edison:2022jln} for recent discussion
and detailed examples of evaluating state sums).  As an example of the
two types of expressions appearing in \cref{eq:SchematicCutEquation},
the iterated two-particle cut can be
represented as a product of trees via
\begin{equation}
  \label{eq:example-cut-prod}
  \ExCutLabeled{1}{2}{3}{4} = \!\!\!\!\! \sum_{\substack{\text{states of }\\ \{\ell_1, \ell_2, \ell_3, \ell_4\}}} \!\!\!\!\!\!\!
  \atree(1,2,\ell_1,\ell_2) \atree(-\ell_1,-\ell_2,\ell_3,\ell_4) \atree(-\ell_3,-\ell_4,3,4)
\end{equation}
or as a collection of diagram numerators via
\begin{align}
  \label{eq:example-cut-sum}
  \ExCutLabeled{1}{2}{3}{4} &=
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}{(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7})^2}
  \ N\left[
    \IUCContrib{1}{4}{3}{2}{\ell_5}{\ell_6}{\ell_7}
                              \right] \notag \\
  &\quad + 
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}
    {(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7'})^2}
  \ N\left[
    \IUCContrib{1}{3}{4}{2}{\ell_5}{\ell_6}{\ell_7'}
    \right] + 25\text{ others.}
\end{align}
When drawing cut diagrams, we will use gray blobs as in the left-hand
sides of \cref{eq:example-cut-prod,eq:example-cut-sum} and all drawn
legs are assumed to be on-shell.  When drawing numerator contributions
to cuts, we will not use blobs on vertices, and will use
\textcolor{cut-color}{dashed lines} bisecting edges to denote cut
lines and use colored edges to highlight the
\textcolor{ucp-color}{uncut propagators}.  From now on we will leave
the cut $\delta(\ell^2)$ factors implicit.

In color-charged theories, both sides of the cut can be further
decomposed according to the color algebra.  For theories charged under
the adjoint of $SU(N)$ via $f^{abc}$-dressed amplitudes, it is
convenient to project onto a specific element of the Del
Duca-Dixon-Maltoni (DDM) color basis \cite{DixonMaltoni}.  This is done by only
inserting color-ordered amplitudes, \emph{a la} the right hand side of
\cref{eq:ddm-basis} in the product of trees, and restricting the set
of compatible diagrams to those whose color factors reduce to the
appropriate DDM element when summing over \emph{only the uncut color
  contractions}.  In this situation, the color becomes an overall
factor on the entire cut equality and thus can be ignored.  Cuts that
are organized in this manner are known as ``color-ordered'' cuts, in
the same sense as color-ordered tree amplitudes.  Since the DDM basis
(and dual Kleiss-Kuijf amplitude basis \cite{Kleiss:1988ne}) have
$(n-2)!$ elements, one generally needs to evaluate
$(n_1-2)! (n_2-2)!...$ separate color-orderings of the same cut, where
$n_1$, $n_2$... are the multiplicities of the amplitudes making up the
cut. As an example, consider the ``bowtie'' cut topology -- one of the
factorization channels of the iterated unitarity cut -- that has two
topologically distinct color-ordered expansions
\begin{equation}
  \label{eq:kt-color-ord}
\!\!\! \!\!\! \!  \NMCutC =
  \begin{cases}
     s\text{-}t \text{ channel}
    &
      \frac{1}{\textcolor{ucp-color}{s_{12}}}
      N \left[\scalebox{0.65}{\KTBTContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      N \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]\\
    t\text{-}u \text{ channel}
    & 
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      N \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
      N\left[
      \scalebox{0.65}{\KTDBContrib{1}{3}{4}{2}}
      \right]
  \end{cases}
\end{equation}
which serve as a complete basis for the full cut.

>>>>

\documentclass[11pt,letter]{article}
\pdfoutput=1
\usepackage{jheppub}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{slashed}
\usepackage{anyfontsize}
\usepackage{tocloft}
\renewcommand{\cftdot}{.}
\renewcommand{\cftsubsecleader}{\cftdotfill{\cftdotsep}}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{bbold}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage[normalem]{ulem}

@@

\author{\large Alex Edison,}
\author{\large James Mangan,}
\author{\large and Nicolas H. Pavao}

@@

\section{Chern-Simons Theory}\label{intro}
The final theory that is unified by semi-abelian Yang-Mills is that Chern-Simons theory. As we will show below, plugging in appropriate on-shell states to the semintegrands constructed in the text will produce precisely the kinematic numerators for this theory. 
\begin{align}
\cubic{hgrey0}{nhpRed}{}{nhpRed}{}{nhpRed}{} = \epsilon^{\mu\nu\rho} \varepsilon^{\mu}_1\varepsilon^{\nu}_2\varepsilon^{\rho}_3 \qquad  &\Rightarrow  \qquad   \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\\
\propFrule{$A^\nu$}{$A^\mu$}{$k \rightarrow $}{nhpRed} = \frac{i}{k^2} \epsilon^{\mu\nu\rho} k^\rho \qquad  &\Rightarrow  \qquad \propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2}= \frac{i}{k^2} \eta^{\mu\nu}
\end{align}
\bibliographystyle{JHEP}
\bibliography{Refs_2loopNLSM}
\end{document}
